Results Analysis

### 2024-04-24 analysis/resultsAnalysis.Rmd
added limits analysis
Now contains all the analysis code for limitations in both journal samples

```{r}
# Load libraries
library(here)
library(lsr)
library(tidyverse)
library(tidylog)
library(binom)
library(ggplot2)
library(RColorBrewer)
```

@@

Remove excluded journals and create separate tibbles for prominent and randomly sampled journals
```{r}
dj_prominent <- d_journals %>%
  filter(sample_id != "random") %>%
  filter(is_empirical == TRUE)

dj_random <- d_journals %>%
  filter(sample_id != "prominent") %>%
  filter(is_empirical == TRUE)
```
## Journal characteristics
Descriptive statistics

@@

Count randomly sampled journals listed as COPE members 
```{r}
randomCOPE <- dj_random %>%
  filter(ppc_type == 'A') %>%
  filter(cope == TRUE) %>%
  summarise(count = n())

@@

prominentPPC
```
*Note: This may change once ARCHIVE ONLY coding is removed*

Inspect post-publication critique types at prominent journals
```{r}
prominentTypes <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(count = n()) %>%
  ungroup()

prominentTypes
```

Inspect how frequently post-publication critique policies were stated for randomly sampled journals 
```{r}
randomPPC <- dj_random %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

@@

Calculate confidence interval for explicitly stated post-publication critique policies among randomly sampled journals
```{r}

@@

randomPPC_CI
```

Inspect post-publication critique types at randomly sampled journals
```{r}
randomTypes <- dj_random %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(count = n()) %>%
  ungroup()

@@

>>>>

```{r}
# Load libraries
library(here)
library(lsr)
library(tidyverse)
library(tidylog)
library(binom)
library(ggplot2)
library(RColorBrewer)
library(assertthat) 
```

@@

Remove excluded journals and create separate tibbles for prominent and randomly selected journals
```{r}
dj_prominent <- d_journals %>%
  filter(sample_id != "random") %>%
  filter(is_empirical == TRUE)

dj_random <- d_journals %>%
  filter(sample_id != "prominent") %>%
  filter(is_empirical == TRUE)
```
```{r}
# Run a test to ensure all randomly-selected and non-empirical journals removed from dj_prominent
assert_that(
  all(dj_prominent$sample_id == "prominent" | dj_prominent$sample_id == "both" 
      & dj_prominent$is_empirical == TRUE),
  msg = "Found unexpected values in the sample_id or is_empirical columns"
  )

# Run a test to ensure all prominent sample and non-empirical journals removed from dj_prominent
assert_that(
  all(dj_random$sample_id == "random" | 
        dj_random$sample_id == "both" 
      & dj_random$is_empirical == TRUE),
  msg = "Found unexpected values in the sample_id or is_empirical columns"
  )

```

## Journal characteristics
Descriptive statistics

@@

Count randomly selected journals listed as COPE members 
```{r}
randomCOPE <- dj_random %>%
  filter(ppc_type == 'A') %>%
  filter(cope == TRUE) %>%
  summarise(count = n())

@@

prominentPPC
```
Get the names of the prominent journals that have 'NO EXPLICIT' PPC options
```{r}
dj_prominent %>%
  filter(ppc_type == "A" & has_ppc == "NO EXPLICIT") %>%
  print()
```
Get how many prominent journals offered more than one type
```{r}
# Return all rows with ppc_name values for ppc_type B and C
dj_prominent %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name)))

# Return count
dj_prominent %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name))) %>%
  summarise(count = n())
```

Inspect post-publication critique types at prominent journals
```{r}
# Counts for all types of PPC, commentary only, and letters only
promAllTypes <- dj_prominent %>%
  filter(!is.na(ppc_name)) %>%
  summarise(n())

promCommentary <- dj_prominent %>%
  filter(ppc_name == "Commentary") %>%
  summarise(n())

promLetters <- dj_prominent %>%
  filter(ppc_name == "Letters") %>%
  summarise(n())

# These numbers will be used in percentage calculations for limits
promAllTypes # Total number of PPC options found in prominent journals
promCommentary # Total number of 'Commentary' options
promLetters # Total number of 'Letters' options
```
Inspect the COPE status of prominent journals with and without PPC policies
```{r}
promCOPEtable <- dj_prominent %>%
 filter(ppc_type == "A") %>%
 mutate(has_ppc = case_when(
    has_ppc == "NO EXPLICIT" ~ "NO",
    has_ppc == "NO IMPLICIT" ~ "NO",
    TRUE ~ has_ppc
 )) %>%
 group_by(has_ppc) %>%
 summarise(
    total_count = n(),
    COPE_count = sum(cope == TRUE)
 ) %>%
 mutate(COPEpercentage = round((COPE_count / total_count) * 100)) %>%
 ungroup()

promCOPEtable
```

Inspect how frequently post-publication critique policies were stated for randomly selected journals 
```{r}
randomPPC <- dj_random %>%
  filter(ppc_type == 'A') %>%
  group_by(has_ppc) %>%
  summarise(count = n()) %>%
  ungroup()

@@

Calculate confidence interval for explicitly stated post-publication critique policies among randomly selected journals
```{r}

@@

randomPPC_CI
```
Get how many randomly selected journals offered more than one type
```{r}
# Return all rows with ppc_name values for ppc_type B and C
dj_random %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name)))

# Return count
dj_random %>%
  filter((ppc_type == "B" | ppc_type == "C") & (!is.na(ppc_name))) %>%
  summarise(count = n())
```

Get the names of the randomly samples journals that have 'NO EXPLICIT' PPC options
```{r}
dj_random %>%
  filter(ppc_type == "A" & has_ppc == "NO EXPLICIT") %>%
  print()
```

Inspect post-publication critique types at randomly selected journals
```{r}
randomTypes <- dj_random %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(count = n()) %>%
  ungroup()

@@

Inspect the COPE status of randomly selected journals with and without PPC policies
```{r}
randCOPEtable <- dj_random %>%
 filter(ppc_type == "A") %>%
 mutate(has_ppc = case_when(
    has_ppc == "NO EXPLICIT" ~ "NO",
    has_ppc == "NO IMPLICIT" ~ "NO",
    TRUE ~ has_ppc
 )) %>%
 group_by(has_ppc) %>%
 summarise(
    total_count = n(),
    COPE_count = sum(cope == TRUE)
 ) %>%
 mutate(COPEpercentage = round((COPE_count / total_count) * 100)) %>%
 ungroup()

randCOPEtable
```

# What limits did journals place on post-publication critique?
Find the most restrictive limits of any journal
```{r}
# Find the lowest length limit
minLength <- d_journals %>%
 select(c(sample_id, journal_name, ppc_name, ppc_length)) %>%
 filter(!is.na(ppc_name),
         ppc_length != "NOT STATED" & ppc_length != "Qualitative") %>%
 mutate(quantLength = as.numeric(ppc_length)) %>%
 slice_min(quantLength, n = 1) %>% # Select the row with the minimum quantLength
 ungroup()

minLength
```

``` {r}
# Find the lowest time-to-submit limit
minTime <- d_journals %>%
 select(c(sample_id, journal_name, ppc_name, ppc_time)) %>%
 filter(!is.na(ppc_name),
         ppc_time != "NOT STATED" & ppc_time != "Qualitative") %>%
 mutate(quantTime = as.numeric(ppc_time)) %>%
 slice_min(quantTime, n = 1) %>% # Select the row with the minimum quantTime
 ungroup()

minTime
```

## Calculations for Table 1: Prominent journals
Check length limits placed on PPC options in prominent journals
```{r}
# Summarise length statistics for All Types
prominentLengthAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_length)) %>%
 mutate(quantLength = case_when(
    ppc_length == "NOT STATED" | ppc_length == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_length) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_length == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_length == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianLength = median(quantLength, na.rm = TRUE),
    Q1 = quantile(quantLength, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantLength, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

prominentLengthAll
```

``` {r}
# Summarise length Statistics for each PPC type
prominentLengthType <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_length)) %>%
 mutate(quantLength = case_when(
    ppc_length == "NOT STATED" | ppc_length == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_length) # create a column that contains only quantitative limits
  )) %>%
 group_by(ppc_name) %>%
 summarise(
    anyLimit = n() - sum(ppc_length == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_length == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianLength = median(quantLength, na.rm = TRUE),
    Q1 = quantile(quantLength, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantLength, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 ) %>%
  ungroup()

prominentLengthType
```
Check time-to-submit limits placed on PPC options in prominent journals
```{r}
# Summarise time-to-submit statistics for All Types
prominentTimeAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_time)) %>%
 mutate(quantTime = case_when(
    ppc_time == "NOT STATED" | ppc_time == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_time) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_time == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_time == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianTime = median(quantTime, na.rm = TRUE),
    Q1 = quantile(quantTime, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantTime, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

prominentTimeAll
```

``` {r}
# Summarise time-to-submit Statistics for each PPC type
prominentTimeType <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_time)) %>%
 mutate(quantTime = case_when(
    ppc_time == "NOT STATED" | ppc_time == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_time) # create a column that contains only quantitative limits
  )) %>%
  group_by(ppc_name) %>%
 summarise(
    anyLimit = n() - sum(ppc_time == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_time == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianTime = median(quantTime, na.rm = TRUE),
    Q1 = quantile(quantTime, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantTime, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 ) %>%
  ungroup()

prominentTimeType
```
Check reference limits placed on PPC options in prominent journals
```{r}
# Summarise reference statistics for All Types
prominentRefAll <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_ref)) %>%
 mutate(quantRef = case_when(
    ppc_ref == "NOT STATED" | ppc_ref == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_ref) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_ref == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_ref == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianRef = median(quantRef, na.rm = TRUE),
    Q1 = quantile(quantRef, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantRef, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

prominentRefAll
```

```{r}
# Summarise reference statistics by type
prominentRefType <- dj_prominent %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_ref)) %>%
 mutate(quantRef = case_when(
    ppc_ref == "NOT STATED" | ppc_ref == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_ref) # create a column that contains only quantitative limits
  )) %>%
  group_by(ppc_name) %>%
 summarise(
    anyLimit = n() - sum(ppc_ref == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_ref == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianRef = median(quantRef, na.rm = TRUE),
    Q1 = quantile(quantRef, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantRef, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 ) %>%
  ungroup()

prominentRefType
```
Check how many PPC options are sent for external peer review in prominent journals
```{r}
# Peer review statistics for All Types
prominentReviewAll <- dj_prominent %>%
  select(c(journal_name, ppc_name, ppc_review)) %>%
  filter(!is.na(ppc_name)) %>%
  summarise(anyReview = n() - sum(ppc_review == "NOT STATED"),
            yesReview = anyReview - sum(ppc_review == "NO"),
            yesPercent = round((yesReview / n()) * 100)   # Percentage reviewed from All Types 
  )

prominentReviewAll

```

```{r}
# Peer review statistics by types
prominentReviewAll <- dj_prominent %>%
  select(c(journal_name, ppc_name, ppc_review)) %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(anyReview = n() - sum(ppc_review == "NOT STATED"),
            yesReview = anyReview - sum(ppc_review == "NO"),
            yesPercent = round((yesReview / n()) * 100)
  ) %>%
  ungroup()

prominentReviewAll

```

## Calculations for Table 2: Randomly selected journals

Check length limits placed on PPC options in randomly selected journals
```{r}
# Summarise length statistics for All Types
randomLengthAll <- dj_random %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_length)) %>%
 mutate(quantLength = case_when(
    ppc_length == "NOT STATED" | ppc_length == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_length) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_length == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_length == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianLength = median(quantLength, na.rm = TRUE),
    Q1 = quantile(quantLength, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantLength, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

randomLengthAll
```

``` {r}
# Summarise length Statistics for each PPC type
randomLengthType <- dj_random %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_length)) %>%
 mutate(quantLength = case_when(
    ppc_length == "NOT STATED" | ppc_length == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_length) # create a column that contains only quantitative limits
  )) %>%
 group_by(ppc_name) %>%
 summarise(
    anyLimit = n() - sum(ppc_length == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_length == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianLength = median(quantLength, na.rm = TRUE),
    Q1 = quantile(quantLength, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantLength, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 ) %>%
  ungroup()

randomLengthType
```
Check time-to-submit limits placed on PPC options in prominent journals
```{r}
# Summarise time-to-submit statistics for All Types
randomTimeAll <- dj_random %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_time)) %>%
 mutate(quantTime = case_when(
    ppc_time == "NOT STATED" | ppc_time == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_time) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_time == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_time == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianTime = median(quantTime, na.rm = TRUE),
    Q1 = quantile(quantTime, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantTime, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

randomTimeAll
```

``` {r}
# Summarise time-to-submit Statistics for each PPC type
randomTimeType <- dj_random %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_time)) %>%
 mutate(quantTime = case_when(
    ppc_time == "NOT STATED" | ppc_time == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_time) # create a column that contains only quantitative limits
  )) %>%
  group_by(ppc_name) %>%
 summarise(
    anyLimit = n() - sum(ppc_time == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_time == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianTime = median(quantTime, na.rm = TRUE),
    Q1 = quantile(quantTime, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantTime, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 ) %>%
  ungroup()

randomTimeType
```
Check reference limits placed on PPC options in prominent journals
```{r}
# Summarise reference statistics for All Types
randomRefAll <- dj_random %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_ref)) %>%
 mutate(quantRef = case_when(
    ppc_ref == "NOT STATED" | ppc_ref == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_ref) # create a column that contains only quantitative limits
  )) %>%
 summarise(
    anyLimit = n() - sum(ppc_ref == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_ref == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianRef = median(quantRef, na.rm = TRUE),
    Q1 = quantile(quantRef, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantRef, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 )

randomRefAll
```

```{r}
# Summarise reference statistics by type
randomRefType <- dj_random %>%
 filter(!is.na(ppc_name)) %>% 
 select(c(journal_name, ppc_name, ppc_ref)) %>%
 mutate(quantRef = case_when(
    ppc_ref == "NOT STATED" | ppc_ref == "Qualitative" ~ NA_real_,
    TRUE ~ as.numeric(ppc_ref) # create a column that contains only quantitative limits
  )) %>%
  group_by(ppc_name) %>%
 summarise(
    anyLimit = n() - sum(ppc_ref == "NOT STATED"),
    anyPercent = round((anyLimit / n()) * 100),
    quantLimit = anyLimit - sum(ppc_ref == "Qualitative"),
    quantPercent = round((quantLimit / n()) * 100),
    medianRef = median(quantRef, na.rm = TRUE),
    Q1 = quantile(quantRef, probs = 0.25, na.rm = TRUE),
    Q3 = quantile(quantRef, probs = 0.75, na.rm = TRUE),
    IQR = Q3 - Q1 
 ) %>%
  ungroup()

randomRefType
```
Check how many PPC options are sent for external peer review in prominent journals
```{r}
# Peer review statistics for All Types
randomReviewAll <- dj_random %>%
  select(c(journal_name, ppc_name, ppc_review)) %>%
  filter(!is.na(ppc_name)) %>%
  summarise(anyReview = n() - sum(ppc_review == "NOT STATED"),
            yesReview = anyReview - sum(ppc_review == "NO"),
            yesPercent = round((yesReview / n()) * 100)   # Percentage reviewed from All Types 
  )

randomReviewAll

```

```{r}
# Peer review statistics by types
randomReviewAll <- dj_random %>%
  select(c(journal_name, ppc_name, ppc_review)) %>%
  filter(!is.na(ppc_name)) %>%
  group_by(ppc_name) %>%
  summarise(anyReview = n() - sum(ppc_review == "NOT STATED"),
            yesReview = anyReview - sum(ppc_review == "NO"),
            yesPercent = round((yesReview / n()) * 100)
  ) %>%
  ungroup()

randomReviewAll

```
